---
title: "HW2-EDA"
author: "Feiran Ji"
date: "4/23/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = TRUE, eval = TRUE}
library(igraph, quietly = TRUE)
library(statnet, quietly = TRUE)
library(Matrix, quietly = TRUE)
```


```{r}
# download dataset
link <- "https://snap.stanford.edu/data/facebook_combined.txt.gz"
download.file(link, destfile = "web-fb.txt.gz")
fb.data <- read.table("web-fb.txt.gz", sep=" ", stringsAsFactors = FALSE)

```


```{r, echo = TRUE, eval = TRUE}
# create igraph from edgelist
fb.igraph <- graph.data.frame(fb.data, directed = FALSE)
# adjacency matrix from igraph
fb.adjacency <- as_adj(fb.igraph)
fb.edgelist <- as.matrix(fb.data) + 1
fb.network <- network(fb.edgelist)
```

## Shortest Paths
First, we will identify the shortest paths are between each pair of nodes. 

```{r, echo = TRUE, eval = TRUE}
shortest.distances <- distances(fb.igraph, mode = "all") 

#network diameter: the longest shortest path that is not infinite
max(shortest.distances[which(shortest.distances < Inf)])

#what is the max shortest path?
max(shortest.distances)

#infinite tells us that there are isolates!
```
There is no isolate account in this network. 


## Network Centrality Measures and Visualization

### Plot the data
```{r}
fb.x <- plot(fb.network, main=paste("Facebook Network"), edge.col = "grey50")
```


### Calculate In-Degree, Out-Degree, Eigenvector, Betweenness, and Closeness Centralities
```{r, echo = TRUE, eval = TRUE}
#in degree centrality
in.degree.centrality <- colSums(as.matrix(fb.adjacency))

#out degree centrality
out.degree.centrality <- rowSums(as.matrix(fb.adjacency))

#eigenvector centrality
eigenvector.centrality <- eigen_centrality(fb.igraph, directed = FALSE)$vector

#normalized betweenness centrality
betweenness.centrality <- estimate_betweenness(fb.igraph, directed = TRUE, cutoff = 10)

#closeness centrality
closeness.centrality <- estimate_closeness(fb.igraph, mode = "total", normalized = FALSE, cutoff = 10)
```



### Plot Network According to Centrality Measures
```{r, echo = TRUE}

#plot in-degree
plot(fb.network, main = paste("In-Degree Centrality"), usearrows = TRUE, vertex.cex = in.degree.centrality / 50, edge.col = "grey50", coord = fb.x)

#plot out-degree
plot(fb.network, main = paste("Out-Degree Centrality"), usearrows = TRUE, vertex.cex = out.degree.centrality / 50, edge.col = "grey50", coord = fb.x)

#plot eigenvector
plot(fb.network, main = paste("Eigenvector Centrality"), usearrows = TRUE, vertex.cex = eigenvector.centrality*4 + 1, edge.col = "grey50", coord = fb.x)

#plot betweenness
plot(fb.network, main = paste("Betweenness Centrality"), usearrows = TRUE, vertex.cex = betweenness.centrality / 15000, edge.col = "grey50", coord = fb.x)

#plot closeness
plot(fb.network, main = paste("Closeness Centrality"), usearrows = TRUE, vertex.cex = closeness.centrality * 150 + 1, edge.col = "grey50", coord = fb.x)
```

## Community Detection
We ran 5 different community detection algorithms on Facebook network and compare the results that we get from each method.

```{r, echo = TRUE, eval = TRUE}

f_g <- cluster_fast_greedy(fb.igraph)

#look at membership
f_g$membership

#look at modularity score
f_g$modularity

#what are the sizes of the class labels?
table(f_g$membership)

#plotting the network according to community label
par(mfrow = c(1,2))
plot(fb.network, coord = fb.x, main = "Original Network")
plot(fb.network, coord = fb.x, main = "Fast and Greedy", vertex.col = f_g$membership)

#plotting the community structure using our own defined colors
plot(fb.network, coord = fb.x, main = "Fast and Greedy", vertex.col = f_g$membership)
```


```{r}
fb.igraph2 = simplify(fb.igraph)
#Running 5 different community detection methods and compare
f_g <- cluster_fast_greedy(fb.igraph2)
info.clusters <- cluster_infomap(fb.igraph2)
l_p <- cluster_label_prop(fb.igraph2)
louvain <- cluster_louvain(fb.igraph2)
walktrap <- cluster_walktrap(fb.igraph2)

#size of communities for each method
table(f_g$membership)
table(info.clusters$membership)
table(l_p$membership)
table(louvain$membership)
table(walktrap$membership)

```

```{r}
par(mfrow = c(2, 3))
plot(fb.network, coord = fb.x, main = "Original")
plot(fb.network, coord = fb.x, main = "Fast and Greedy",
     vertex.col = f_g$membership)
plot(fb.network.clusters$membership)
plot(fb.network, coord = fb.x, main = "Label Propagation", 
     vertex.col = l_p$membership)
plot(fb.network, coord = fb.x, main = "Louvain", 
     vertex.col = louvain$membership)
plot(fb.network, coord = fb.x, main = "Walktrap", 
vertex.col = walktrap$membership)
```

