---
title: "Comp_stat_HW2_PartII"
author: "Jade Yun"
date: "4/22/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(igraph)
library(statnet)
library(Matrix)
library(datasets)
```

1(a)

```{r}
# loading iris 
data(iris)
summary(iris)
```

```{r}
# Separate data from class labels
irisData=iris[,1:4]
irisClass=as.numeric(iris$Species)
```


```{r}
# Column standardize iris data to have mean 0 and unit variance
irisData=scale(irisData)
```

```{r}
# Calculate distance matrix between all pairs of flowers
irisDist=as.matrix(dist(irisData))
```

```{r}
forHist=irisDist[lower.tri(irisDist,diag=FALSE)]
quartz()
hist(forHist)
```

```{r}
# 25% percentile is of distances.
quart25=quantile(forHist,.25)
```

```{r}
# Create an adjacency matrix template
AdjMat=irisDist
```

```{r}
# make the graph more sparse， remove any edges that are not in the bottom 25% of distances.
AdjMat[AdjMat>quart25]=0
```

```{r}
graph=graph.adjacency(AdjMat,mode='undirecte')
```

```{r}
colorvec=c('lightslateblue','magenta','chartreuse')
V(graph)$color=colorvec[irisClass]
quartz()
plot(graph,vertex.size=6,vertex.label=NA)
```

1(c)

```{r}
plot(graph, layout = layout.circle)
```


```{r}
plot(graph, layout = layout.kamada.kawai)
```


```{r}
plot(graph, layout = layout.fruchterman.reingold)
```

1(d)
IDK !!!!!! >.< !!!!!!


2(a)
```{r}
link <- "https://snap.stanford.edu/data/facebook_combined.txt.gz"
download.file(link, destfile = "web-fb.txt.gz")
snap.data <- read.table("web-fb.txt.gz", sep=" ", stringsAsFactors = FALSE)
```


```{r pressure}
# convert to 'igraph' object 
g <- graph.data.frame(snap.data, directed = FALSE)
```

what do edges and nodes represent?

This dataset is used to analyze users relationships at Facebook. Nodes represent each user and a undirected edge from node i to node j represents that user i is friend with user j on Facebook. 

2(b)

```{r}
vcount(g)
ecount(g)
```

how many vertices and edges are there in the data set?

There are 4039 nodes and 88234 edges. 

2(c)

```{r}
is_weighted(g)
```

Is the network weighted? If so, what do the weights represent?

No, the network is not weighted. 


```{r}
is_directed(g)
```

Is the network directed or undirected?

No, the network is NOT directed. 


2(e)

How much storage will it require to store the network using sparse representation? How about as an adjacency matrix?

Sparse representation: O(E) and Matrix representation: O(V^2)

2(f)

Which user has the most friends? 
Do most users have no more than 200 friends?

2(g)

```{r}
# extract 3 subgraphs, 200 random nodes each

sub.g1 <- induced.subgraph(g, sample(V(g), 200))
sub.g2 <- induced.subgraph(g, sample(V(g), 200))
sub.g3 <- induced.subgraph(g, sample(V(g), 200))
```

```{r}
par(mfrow = c(3, 3))
#igraph_options(vertex.label=NA)
```

```{r}
V(sub.g1)$frame.color <- "white"
V(sub.g1)$color <- "orange"
V(sub.g1)$label <- ""
V(sub.g1)$size <- 5
E(sub.g1)$arrow.mode <- 0
```

```{r}
# subgraph 1 - circle
l <- layout.circle(sub.g1)
plot(sub.g1, layout = l)
```

```{r}
# subgraph 2- Kamada-Kawai
plot(sub.g1, layout = layout.kamada.kawai)
```

```{r}
# subgraph 2- Kamada-Kawai
plot(sub.g1, layout = layout.fruchterman.reingold)
```


```{r}
V(sub.g2)$frame.color <- "white"
V(sub.g2)$color <- "orange"
V(sub.g2)$label <- ""
V(sub.g2)$size <- 5
E(sub.g2)$arrow.mode <- 0
```


```{r}
# subgraph 2 - circle
plot(sub.g2, layout = layout.circle)
```

```{r}
# subgraph 2- Kamada-Kawai
plot(sub.g2, layout = layout.kamada.kawai)
```

```{r}
# subgraph 2 - Fructerman Reingold
plot(sub.g2, layout = layout.fruchterman.reingold)
```

```{r}
V(sub.g3)$frame.color <- "white"
V(sub.g3)$color <- "orange"
V(sub.g3)$label <- ""
V(sub.g3)$size <- 5
E(sub.g3)$arrow.mode <- 0
```

```{r}
# subgraph 3 - circle
plot(sub.g3, layout = layout.circle)
```

```{r}
# subgraph 3 - Kamada-Kawai
plot(sub.g3, layout = layout.kamada.kawai)
```

```{r}
# subgraph 3 - Fructerman Reingold
plot(sub.g3, layout = layout.fruchterman.reingold)
```


Are there any interesting network properties that become apparent when coloring or resizing the vertices?

We could resize the vertices based on degree or centrality and color those has more than 50 friends. 













