---
title: "Comp_stat_HW2_PartII"
author: "Jade Yun"
date: "4/22/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(igraph)
library(statnet)
library(Matrix)
library(datasets)
```

1(a)

```{r}
# loading iris 
data(iris)
summary(iris)
```


2(a)
```{r}
link <- "https://snap.stanford.edu/data/web-Google.txt.gz"
download.file(link, destfile = "web-Google.txt.gz")
snap.data <- read.table("web-Google.txt.gz", sep="\t", stringsAsFactors = FALSE)
```


```{r pressure}
# convert to 'igraph' object 
g <- graph.data.frame(snap.data, directed = TRUE)
```

what do edges and nodes represent?

This dataset is used to analyze relationship between google's web pages. Nodes represent web pages and a directed edge 
from node i to node j represents that page i hyperlinked page j. 

2(b)

```{r}
vcount(g)
ecount(g)
```

how many vertices and edges are there in the data set?

There are 875713 nodes and 5105039 edges. 

2(c)

```{r}
is_weighted(g)
```

Is the network weighted? If so, what do the weights represent?

No, the network is not weighted. 


```{r}
is_directed(g)
```

Is the network directed or undirected?

Yes, the network is directed. 


2(e)

How much storage will it require to store the network using sparse representation? How about as an adjacency matrix?

Sparse representation: O(E) and Matrix representation: O(V^2)

2(f)

Which page contains the most hyperlinks to other pages? 
Which page has been hyperlinked the most by other pages?

2(g)

```{r}
# extract 3 subgraphs, 200 random nodes each

sub.g1 <- induced.subgraph(g, sample(V(g), 200))
sub.g2 <- induced.subgraph(g, sample(V(g), 200))
sub.g3 <- induced.subgraph(g, sample(V(g), 200))
```

```{r}
par(mfrow = c(3, 3))
igraph_options(vertex.label=NA)
```

```{r}
# - SUBGRAPH 1 - 

plot(sub.g1, layout = layout.circle)
```

```{r}
plot(sub.g1, layout = layout.kamada.kawai)
```

```{r}
plot(sub.g1, layout = layout.fruchterman.reingold)
```


```{r}
# - SUBGRAPH 2 - 
plot(sub.g2, layout = layout.circle)
```

```{r}
plot(sub.g2, layout = layout.kamada.kawai)
```

```{r}
plot(sub.g2, layout = layout.fruchterman.reingold)
```

```{r}
# - SUBGRAPH 3 - 
plot(sub.g3, layout = layout.circle)
```

```{r}
plot(sub.g3, layout = layout.kamada.kawai)
```

```{r}
plot(sub.g3, layout = layout.fruchterman.reingold)
```


Are there any interesting network properties that become apparent when coloring or resizing the vertices?

We could resize the vertices based on degree or centrality and color those has been hyperlinked more than 500 times.


Here we have to resize the vertices and color the nodes. I am not sure how to do that..






